# How to Automatically Know if your Gas Mileage needs a Manual Takeover
#### An exploration into the difference in miles per gallon of manaul vs automatic transmissions

### Summary
Blah blah blah
blah blah blah
blah blah blah

### Which transmission type is better for miles per gallon?
The age old struggle of which transmission type is better, automatic or manual, is an age old struggle that we here at *Motor Trend* do not take lightly.  We are here to take you on a journey examining which transmission is ultimately the best for getting the most from your dollar per mile traveled.  Get ready to dive in becasue you are in for a ride!


```{r, echo=FALSE}
library(ggplot2)
library(lattice)

data(mtcars)
data <- mtcars
data$am[data$am == 0] <- "automatic"
data$am[data$am == 1] <- "manual"
auto <- data[data$am == "automatic",]
manual <- data[data$am == "manual",]


ggplot(data, aes(x=mpg, color=am)) + geom_density() + scale_colour_discrete(name = "Transmission Type") + ggtitle("Figure 1. Density plot of MPG by transmission type")
```

```{r}

```

```{r}

```

```{r}

```
